Keep your Laziness in Check

### 2018-02-14 paper/main.tex
[paper] Restructure after today's discussion

% !!! PLEASE DON'T CHANGE THESE !!! INSTEAD DEFINE YOUR OWN texdirectives.tex !!!
% These flags are for the final camera ready version for POPL
% Draft, full version flags
\newif\ifdraft\drafttrue
\newif\iffull\fullfalse
\newif\ifanon\anonfalse
\newif\iflast\lasttrue

@@

Lazy evaluation gives tremendous power to functional programmers
~\cite{WhyFPMatters} and is actively used in various practical
applications. For instance, the big-data framework Spark uses laziness
to drive computations on demand\cn, Apple's programming language Swift
leverages it to initialize expensive resources as needed\cn, and many
functional reactive programming web frameworks crucially rely on
it\cn \leo{This point might need more exposition about frp}. Haskell
is the most popular programming language that allows programmers to
enjoy the power of lazy evaluation by default. However, that power
does not come without a price: reasoning about lazy programs is
notoriously hard. Let us consider two examples that demonstrate the
power of laziness, as well as the intricacies of programming with it.

@@

As these examples show, the power of laziness puts great
responsibility on programmers' shoulders. An incorrect implementation
can either behave functionally the same but with inferior performance,
or it can completely break down! Reasoning about lazy programs is also
notoriously hard --- the programmer needs to have a mental model of the
order of evaluation of the code, and this order is not always obvious
from the syntax of the code.

\leo{Didn't we have an FRP quote by someone that Laziness was the third source of
bugs or something? Work that in?}
\hzh{Ryan Trinkle (author of Reflex) said laziness was the 3rd most difficult thing to
get right, but let's check with him on whether we can quote him on that.}

How, then, can we test and catch these laziness bugs during development? Is it
possible to precisely specify laziness behavior just as we would for functional
correctness?
%
In this paper, we demonstrate techniques for observing laziness in Haskell, and
propose a system of creating specifications of lazy Haskell programs. We
implement these ideas as a Haskell library.

Our technical contributions are as follows:
\begin{itemize}
\item We introduce a novel approach to specifying the laziness behavior of functions that is
{\em precise} \hzh{and compositional}. Such specifications are
expressive enough to capture exactly how much evaluation occurs within
a data structure \leo{Also add drawbacks/partiality here? If not here,
where in the intro?} \hzh{I'm not sure if we should bring in
partiality (in the sense of divergence/nontermination) yet since that
seems like a rather technical detail of the spec}
\item We propose a framework for testing against such specifications.
\item We improve upon the random function generators from QuickCheck to generate functions with arbitrary laziness behaviors.
\item We demonstrate non-trivial generic programming techniques to apply StrictCheck to almost all data types.
\end{itemize}
%
Section~\ref{sec:related} discusses related work. We conclude and draw
directions for future work in Section~\ref{sec:concl}.

>>>>

% !!! Please Don't CHANGE THESE !!! INSTEAD DEFINE YOUR OWN texdirectives.tex !!!
% These flags are for the final camera ready version for POPL
% Draft, full version flags
\newif\ifdraft\drafttrue
\newif\iffull\fullfalse
\newif\ifanon\anonfalse
\newif\iflast\lasttrue

@@

Lazy evaluation gives tremendous power to functional programmers
~\cite{WhyFPMatters} and is actively used in various practical
applications. For instance, the big-data framework Spark uses laziness
to drive computations on demand\cn, Apple's programming language Swift
leverages it to initialize expensive resources as needed\cn, and many
functional reactive programming web frameworks crucially rely on
it\cn \leo{This point might need more exposition about frp}. Haskell
is the most popular programming language that allows programmers to
enjoy the power of lazy evaluation by default. However, that power
does not come without a price.
\leo{Didn't we have an FRP quote by someone that Laziness was the third source of
bugs or something? Work that in?}
\hzh{Ryan Trinkle (author of Reflex) said laziness was the 3rd most difficult thing to
get right, but let's check with him on whether we can quote him on that.}

The power of laziness puts great responsibility on programmers'
shoulders. An incorrect implementation can either behave functionally
the same but with inferior performance, or it can completely break
down! Reasoning about lazy programs is also notoriously hard --- the
programmer needs to have a mental model of the order of evaluation of
the code, and this order is not always obvious from the syntax of the
code.

How, then, can we test and catch these laziness bugs during
development? Is it possible to precisely specify laziness behavior
just as we would for functional correctness?
%
In this paper, we demonstrate techniques for observing laziness in
Haskell, and propose a system of creating specifications of lazy
Haskell programs. We implement these ideas as a Haskell library. For
concreteness, consider the following small example.

\leo{Insert take example here. There is enough space to have both code
and contributions in this page. yay :) }

In the next section we work through two more examples, functional
queues and functional maps, demonstrating both the difficulties of
programming with laziness and how our approach can be used to
alleviate them.
%
In the rest of the paper, we present our technical contributions:
\begin{itemize}
\item We introduce a novel approach to specifying the laziness behavior
of functions that is {\em precise} and {\em compositional}. Such
specifications are expressive enough to capture exactly how much
evaluation occurs within a data structure
%\leo{Also add
%drawbacks/partiality here? If not here, where in the intro?} \hzh{I'm
%not sure if we should bring in partiality (in the sense of
%divergence/nontermination) yet since that seems like a rather
%technical detail of the spec}
\item We propose a framework for testing against such specifications.
\item We improve upon the random function generators from QuickCheck to generate functions with arbitrary laziness behaviors.
\item We demonstrate non-trivial generic programming techniques to apply StrictCheck to almost all data types.
\end{itemize}
%
Section~\ref{sec:related} discusses related work. We conclude and draw
directions for future work in Section~\ref{sec:concl}.

\leo{These examples should go in the next section. Putting them here to avoid conflicts with Hengchu}

@@
